# Results



```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library (tidyverse)
library(lubridate)
library(parcoords)
library(plotly)
library(vcd)

citations = read.csv("citations.csv")

citations_count = citations%>%count(cited_doi)

```

## Frequency Distribution of Citations received by papers
One of the major networking or social aspect of the academic research community is authors writing papers and citing other author's work. But as the corpus of research articles is huge, very few authors and their papers get a very high number of citations for their work. Sidney Redner in her paper, Citation Statistics from 110 Years of Physical Review - http://physics.bu.edu/~redner/pubs/pdf/PT.pdf - calls out this somewhat alarming statistical phenomenon as follows - "A somewhat depressing observation is that nearly 70% of all PR articles have been cited fewer than 10 times. Also evident is the small number of highly cited publications" 

In our dataset, we have 563557 papers, with a median citations received value as 6, and a mean citations received value ~15, which is not very high.
```{r echo=FALSE, warning=FALSE, message=FALSE}
summary(citations_count)

```

The histogram further confirms that the data is heavily right skewed. A very few papers receive more than 20 citations as shown below.  (We choose number of citations received <50, as there are very few papers in each bucket more than 50 citations)

```{r echo=FALSE, warning=FALSE, message=FALSE}
main <- subset(citations_count,n < 50)

ggplot(data=main, aes(x = n)) +
  geom_histogram(binwidth = 5) +
  ggtitle("Histogram of Number of Citations (n < 50)")+
  xlab("n (Number of citations received by papers)")+
  ylab("Count of papers that have been cited n times")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))

```

We also plot the histogram of papers which have citations received between 50 and 500 separately to show that the y-scale is different in this graph by several orders of magnitude compared to the above graph.  
Together the above and below graph cover almost 99% of the papers in our data set. The change in scale indicates the sharp power law relationship between the number of papers receiving a given number of citations.  

```{r echo=FALSE, warning=FALSE, message=FALSE}
main_1 <- subset(citations_count,n >= 50 & n<=500)

ggplot(data=main_1, aes(x = n)) +
  geom_histogram(binwidth = 5) +
  ggtitle("Histogram of Number of Citations (50<=n<=500)")+
  xlab("n (Number of citations received by papers)")+
  ylab("Count of papers that have been cited n times")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))

```  

## Are there any frequency distributions associated with journal categories?

The different categories of Physical revire papers,based on the topics they cover, have been defined in the section on data sources.  
```{r echo=FALSE, warning=FALSE, message=FALSE}
metadata = read.csv("phpaper_metadata_v2.csv")

metadata$paper_pub_date <- as.Date(metadata$paper_pub_date)

total = merge(x = citations_count, y = metadata, by.x = "cited_doi", by.y = "paper_doi")

total$paper_pub_date <- as.Date(total$paper_pub_date)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data=total, aes(x = n)) +
  geom_histogram() +
  ggtitle("Number of Citations Received by Category") +
  xlab("n (Number of citations received by papers)")+
  ylab("Count of papers that have been cited n times")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))+
  facet_wrap(~paper_journal_category, scales="free")

```

From the chart above, we can see that the citation received histogram of all categories are right skewed. However, the number of citations is more concentrated around zero in some categories, such as PR, PRA, PRB, PRC and PRD while less concentrated in other categories such as PRAB, PRFLUIDS and PRAER.

## Time series analysis of physical review papers
After exploring the data around citations received across all articles, we look at the evolution of publications and citations over time. 

First, we look at the publications data over time to see if there exists any long term secular patterns.

```{r echo=FALSE, warning=FALSE, message=FALSE}
pub_yearly <- metadata %>% group_by(Year = year(paper_pub_date)) %>% summarize(paper_doi = n())

plot_ly(
          pub_yearly, x = ~Year, y = ~paper_doi,
          type = 'scatter',
          mode = 'lines+markers',
          hovertemplate = "%{x}: %{y:.0f}<extra></extra>") %>% layout(title = "Number of Publications by Year", yaxis=list(title=""))

```

Not surprisingly, the number of publications has increased with time. However, the trend is not entirely upward. The number of publications declined from 2012 to 2015 and then recovered from there. 

We are curious whether the number of citations received by papers will exhibit a similar pattern as the number of publications.

```{r echo=FALSE, warning=FALSE, message=FALSE}
total$paper_pub_date <- as.Date(total$paper_pub_date)

yearly <- total %>% 
  group_by(Year = year(paper_pub_date))%>%
  summarize(total_n = sum(n))

plot_ly(
          yearly, x = ~Year, y = ~total_n,
          type = 'scatter',
          mode = 'lines+markers',
          hovertemplate = "%{x}: %{y:.0f}<extra></extra>") %>% layout(title = "Number of Citations Received by Publication Year", yaxis=list(title=""))

```

Based on our data (1893-2019), the number of citations generally went up over time and reached the peak in 2008, which means that paper published in 2008 received the highest number of citations. We think it makes sense to see an inflection point since paper published too long ago might be out of date while paper published in recent years have not had enough time to be cited by other papers. It is interesting to see that the year 2008 is the peak, indicating that on average it takes about 12 years for a paper to receive the highest number of citations.

To further explore, we faceted on months from January to December to see if there exists any seasonal patterns.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10}
monthly <- total %>% 
  group_by(Year = year(paper_pub_date), Month = month(paper_pub_date)) %>% 
  summarize(total_n = sum(n))

ggplot(monthly, aes(Year, total_n)) +
  geom_line() +
  facet_grid(~Month) +
  ylab("") +
  ggtitle("Number of Citations Received by Published Month")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))

```

While we see a similar pattern across all months in the chart above, instances of extremely high citations were associated with papers published in October, March and July.

After exploring the number of citations a paper received over time, we looked into how long on average it takes for a paper to receive a citation.

```{r echo=FALSE, warning=FALSE, message=FALSE}
citations$citing_paper_pub_date=metadata$paper_pub_date[match(citations$citing_doi,metadata$paper_doi)]

citations$cited_paper_pub_date=metadata$paper_pub_date[match(citations$cited_doi,metadata$paper_doi)]

citations$diff = as.numeric(difftime(citations$citing_paper_pub_date, citations$cited_paper_pub_date, units ="days"))/(365.25)

cleancitations <- citations[complete.cases(citations), ] #remove NA

yearly2 <- cleancitations %>% 
  group_by(Year = year(cited_paper_pub_date)) %>%
  summarise(avg_diff = mean(diff))

plot_ly(
  yearly2, x = ~Year, y = ~avg_diff,
  type = 'scatter',
  mode = 'lines+markers',
  hovertemplate = "%{x}: %{y:.0f}<extra></extra>") %>% layout(title = "Average number of years for a paper to be cited by its publication year", yaxis=list(title=""))

```


Interestingly but perhaps not too surprisingly, the average time it takes for a physics paper to be cited by other physics papers generally goes down as times goes by. The downward trend has been especially clear since 1980.

The growing popularity in social media has enabled researchers to promote their work more efficiently, such as opening a Facebook page to encourage casual discussion of their papers or opening a Twitter account and tweet about their paper. Moreover, research shows that tweeting results in significantly more article citations over time, highlighting the durable scholarly impact of social media activity. (https://www.sciencedirect.com/science/article/pii/S0003497520308602)


## A deep-dive into the year 2008
In order to uncover interesting points of our data, we do a deep dive into and a detailed analysis for Year 2008

We are interested in studying the physics papers published in 2008 when the number of citations reached record high.

To start with, we used interactive parallel coordinates to identify outliers and the relationship between different categorical variables, such as journal categories, paper length, number of authors and number of citations.

```{r echo=FALSE, warning=FALSE, message=FALSE}
recent <- total %>% select(paper_pub_date, paper_journal_category, paper_authors_count, paper_length, n) %>% filter(paper_pub_date >= as.Date("2008-01-01") & paper_pub_date <= as.Date("2008-12-31"))

recent %>% select(paper_journal_category, paper_authors_count, paper_length, n) %>%
          parcoords(
                  rownames = F
                  , brushMode = "1D-axes"
                  , reorderable = T
                  , queue = T
                  , alpha = .5)
```

The parallel coordinates plot shows us some interesting trends. First, in contrast to general belief, number of authors is negatively correlated with paper length. Second, most journal categories have author counts spread from 1 to 25 while two categories - RMP and PRSTPER have relatively fewer authors. We suspect this is because these two categories are review articles of other physics papers and therefore have few authors associated with them.

A few outliers jump out from the graph. For instance, one paper in RMP has the highest paper length but was written by fewer than 5 people and received very few citations. Moreover, the paper that received the highest number of citations and relatively high paper length has fewer than 5 authors.

Furthermore, we would like to see how the number of citations varies across journal categories for the publications in 2008. We focus on the following categories that have meaningful number of total citations in 2008: PRA, PRB, PRC, PRD, PRE and PRL. (The definition of each category is shown on page "Data sources".)

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=7}
recent_mo <- as.data.frame(recent)

recent_mo <- filter(recent_mo,paper_journal_category %in% c('PRA','PRB','PRC','PRD','PRE','PRL'))

recent_mo$n <- cut(recent_mo$n, breaks = c(-Inf, 6, 15, Inf), labels = c('<6', '[6,15)','>=15'), right = FALSE) #left closed

mosaic(n~paper_journal_category,direction=c("v","h"),recent_mo, main="Number of Citations Received by Journal Category")
```

Note: We picked 6 and 15 to be the breaks so that the number of citations for each interval will be meaningful  across all categories.

```{r echo=FALSE, warning=FALSE, message=FALSE}
summary(recent_mo$n)
```

The mosaic plot above shows that most papers published in 2008 were in PRB and RPL and that papers in PRL, PRC and PRD received much higher number of citations than PRE, since the percentage of papers with more than 6 citations was much higher in RPL, PRC and PRD than PRE. (The definition of each category is shown on page "Data sources".)

All in all, we were able to see some interesting trends in 2008.

## What helps (or does not help) papers to be cited?
Lastly, we do a deep dive on some factors that contribute to the citations received by papers across all the years in our dataset (1893-2019).
Note: There are outliers in the data viz. very few papers have very high citations, very few papers have more than 25 authors, very few papers are more than 50 pages long, etc. Hence, we choose median for our calculations below as we do not want the outliers to skew the measures.  


### What role do the attributes of a paper - length of the paper, and length of the title of the paper - play? 

1. Length of the title of a paper - An article in the Nature magazine claims that papers with shorter titles get more citations - https://www.nature.com/news/papers-with-shorter-titles-get-more-citations-1.18246. We wanted to see if this was applicable to articles/papers published in the physical review journal. We plot the average number of citations received by papers with a given title length.   
As shown in the scatter plot below, we do see a decreasing trend in the median number of citations as the length of the title of the paper becomes longer. 

```{r echo = FALSE, warning = FALSE, message=FALSE}
metadata_titlecleanup <- metadata[!grepl("<math",metadata$paper_title),]
metadata_titlecleanup$paper_title_length <-sapply(strsplit(metadata_titlecleanup$paper_title, " "), length)
metadata_titlecleanup$n_citations=citations_count$n[match(metadata_titlecleanup$paper_doi,citations_count$cited_doi)]
```

```{r echo = FALSE, warning = FALSE, message=FALSE, fig.width=7, fig.height=5}
ggplot(metadata_titlecleanup, aes(x=paper_title_length, y = n_citations))+
  #stat_summary(fun = "mean", geom="point", aes(color="Mean"))+
  stat_summary(fun = "median", geom="point", color ="blue")+
  xlab("n (Length of the paper title - number of words in the title) ")+
  ylab("Median citations received by a paper of length n")+
  labs(title="Citations received vs Paper title length")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))
```
  
Note: For the above calculation, we remove the papers where titles have a math function in them. This is because the title length includes html references in them making the title length longer than the actual title.  25% of papers in our dataset have such titles, and we exclude them from this calculation only. We believe that 75% of our dataset is a good representation of the entire dataset of 656622 observations.   

```{r echo = FALSE, warning = FALSE, message=FALSE}
metadata[,'paper_authors_count']<-factor(metadata[,'paper_authors_count'])
metadata<-metadata%>%filter(paper_authors_count!=0)
metadata$num_authors <- forcats::fct_collapse(factor(metadata$paper_authors_count), "1"="1", "2" = "2", "3"="3","4"="4","5"="5",other_level = "5+")
```

```{r echo = FALSE, warning = FALSE, message=FALSE}
metadata$n_citations=citations_count$n[match(metadata$paper_doi,citations_count$cited_doi)]
num_citations <- cut(metadata$n_citations, breaks = c(0,25,50,75,100,500,10763), labels = c('<25', '25-50','50-75','75-100','100-500','>500'))
metadata$num_citations <- num_citations
metadata <- metadata %>% mutate(num_citations = fct_rev(num_citations))
```  

```{r echo = FALSE, warning = FALSE, message=FALSE}
metadata_less_than_500cit <- metadata%>%filter(n_citations <500)
```


```{r echo = FALSE, warning = FALSE, message=FALSE}
df2 <- aggregate(x=metadata_less_than_500cit$n_citations, list(metadata_less_than_500cit$paper_length),FUN="median")
names(df2)<-c("paper_length","median_citations")
```

2. Length of the paper - A study by Tahatman, Afshar, Ahamdzadeh on papers published WoS, Scopus, PubMed and Medline believes that longer papers get more citations because they contain more information. (Factors affecting number of citations: a comprehensive review of the literature- https://link.springer.com/article/10.1007/s11192-016-1889-2). 
We observe this phenomenon in physical review journal papers shown in the scatter diagram below.  

```{r echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=4}
g_citation_length<- ggplot(metadata_less_than_500cit, aes(x=paper_length,y=n_citations))
g_citation_length<-g_citation_length + stat_summary(fun = "median", geom="point")
g_citation_length<-g_citation_length+geom_smooth(data = df2, aes(x=paper_length, y = median_citations), span=0.5, fill ="blue" )
g_citation_length + 
  xlab("Length of journal paper (number of pages)") + 
  ylab("Median citations received by the paper")+
  labs(title="Citations received vs paper length") +theme_bw()+theme(plot.title = element_text(hjust=0.5))
```
Note: For the above calculation, we look at papers that have receieved <=500 total citations, as very few papers (less than 15% of our data set) have very high citations. We do this to reduce noise in our data.  

## Are there any social aspects at play?
Now we look at the people or social aspect of citations. We want to see what social factors, if any, play a role in increasing or decreasing citations received.
### Authors of a paper - More the merrier? 
First, we look at whether or not the number of authors of an article has an effect on the number of citations received by a paper. The scatter diagram below shows the median number of citations received by papers that have a given number of authors. 
We believe collaboration/networking helps in boosting the citations of a paper. Having a co-author is always helpful than working alone, but it looks like when the number of authors increases beyond 3, the median citations stay flat until 15 authors, and then increase again and go downward beyond ~18 co-authors. 
As we have data of only physics paper citing physics paper, there might be much more citations received by the paper from other journals like IEEE, etc. We believe that an overall number of citations could indicate a stronger correlation between citations and number of authors, but we are not certain about it.

```{r echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=4}
citations_count$n_authors=metadata$paper_authors_count[match(citations_count$cited_doi,metadata$paper_doi)]
citations_count <- citations_count[complete.cases(citations_count), ]
```

```{r echo = FALSE, warning = FALSE, message=FALSE}
df4 <- aggregate(x=citations_count$n, list(citations_count$n_authors),FUN="median")
names(df4)<-c("n_authors","median_citations")
```



```{r echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=4}
g1<-ggplot(citations_count, aes(x=as.numeric(n_authors), y = n))
g1<- g1+stat_summary(fun = "median", geom="point")
g1<- g1+geom_smooth(data=df4,aes(x=as.numeric(n_authors),y=median_citations), span = 0.5, fill="blue")
  g1+ xlab("n - Number of authors of a paper")+
  ylab("Median number of citations received by a paper with n authors")+
  labs(title="Citations received vs number of authors")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))
```

A paper entitled "The relationship between the number of authors of a publication, its citations and the impact factor of the publishing journal: Evidence from Italy" -https://www.sciencedirect.com/science/article/pii/S1751157715200557 - mentions that the review articles in the field of physics show a high positive correlation between the number of citations received and the number of authors.  
From the scatter plot below, that contains only RMP (Review of Modern Physics) papers,we see that the median citations is higher than the median citations compouted across all categories in the graph above. And even though a high positive correlation is not evident in the below graph, the inclusion of citations outside of physics review journals might provide a clearer evidence.
```{r echo = FALSE, warning = FALSE, message=FALSE}
citations_count$journal_category<- metadata$paper_journal_category[match(citations_count$cited_doi,metadata$paper_doi)]
```


```{r echo = FALSE, warning = FALSE, message=FALSE}
citations_count_RMP<-citations_count%>%filter(citations_count$journal_category=='RMP')
```
`
```{r echo = FALSE, warning = FALSE, message=FALSE}
df5 <- aggregate(x=citations_count_RMP$n, list(citations_count_RMP$n_authors),FUN="median")
names(df5)<-c("n_authors","median_citations")
```



```{r echo = FALSE, warning = FALSE, message=FALSE, fig.width=6, fig.height=4}
g2<-ggplot(citations_count_RMP, aes(x=as.numeric(n_authors), y = n))
g2<- g2+stat_summary(fun = "median", geom="point")
g2<- g2+geom_smooth(data=df5,aes(x=as.numeric(n_authors),y=median_citations), span = 0.5, fill="blue")
  g2+ xlab("n - Number of authors of a paper")+
  ylab("Median number of citations received by a paper with n authors")+
  labs(title="Citations received vs number of authors - RMP papers")+
  theme_bw()+theme(plot.title = element_text(hjust=0.5))
```

Thus we are able to study some interesting factors in play - social factors (authors) and quality of a paper(length and title) - that affect the citations received by papers.
In the next section, we look at how does the citation count of most cited physics papers of nobel prize winners evolve with time, and what happens before and after their nobel prize year. 