---
title: "finalproject_draft"
author: "Akanxa Padhi"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
#Packages
```{r}
library(dplyr)
library(ggplot2)
```

#Code 
```{r}
df_prpaper = read.csv('/Users/akanxa/PycharmProjects/PhysicalReviewData/phpaper_metadata_latest.csv')
df_prpaper$paper_pub_date = as.Date(df_prpaper$paper_pub_date)
df_prpaper$paper_pub_year=format(df_prpaper$paper_pub_date, format = "%Y")
df_papercount_year = df_prpaper%>%count(paper_pub_year)
df_papercount_year$paper_pub_year = as.integer(df_papercount_year$paper_pub_year)

```
```{r}
g_pub_time <- ggplot(df_papercount_year, aes(x=paper_pub_year, y = n))+
  geom_point()

g_pub_time + facet_wrap(.~(df_papercount_year$paper_pub_year - (df_papercount_year$paper_pub_year %% 10)), scales = 'free', ncol=3, nrow = 6)
```

```{r}
df_prpapercitations = read.csv('/Users/akanxa/PycharmProjects/PhysicalReviewData/aps-dataset-citations-2019.csv')
df_citationsreceived = df_prpapercitations%>%count(cited_doi)
```

```{r}
#citation_count <-cut(df_lakes$Species,breaks = 3,labels = c('small','medium','large'))
g <- ggplot(df_citationsreceived, aes(x=n)) +
  geom_density()
plot(g)
```

```{r}
df_prpaper%>%select(paper_doi, paper_pub_date)%>%filter(grepl('Clifford G. Shull',df_prpaper$paper_authors))
df_citationsreceived%>%filter(cited_doi=='10.1103/RevModPhys.67.753')
```
```{r}
df_prpapercitations$citing_paper_pub_date=df_prpaper$paper_pub_date[match(df_prpapercitations$citing_doi,df_prpaper$paper_doi)]
df_prpapercitations$cited_paper_pub_date=df_prpaper$paper_pub_date[match(df_prpapercitations$cited_doi,df_prpaper$paper_doi)]
```
```{r}
df_citationsreceived$n_authors=df_prpaper$paper_authors_count[match(df_citationsreceived$cited_doi,df_prpaper$paper_doi)]
ggplot(df_citationsreceived, aes(x=factor(n_authors), y = n))+stat_summary(fun.y = "mean", geom="point")+theme_light()
```
```{r}
df_temp <- df_prpaper[!grepl("<math",df_prpaper$paper_title),]
df_temp$paper_title_length <-sapply(strsplit(df_temp$paper_title, " "), length)
df_temp$n_citations=df_citationsreceived$n[match(df_temp$paper_doi,df_citationsreceived$cited_doi)]
ggplot(df_temp, aes(x=paper_title_length, y = n_citations))+stat_summary(fun.y = "mean", geom="point")+theme_light()
```
```{r}
#citation_count <- cut(df_citationsreceived$n, breaks = 5, labels = c('very few citations', 'few #citations','average citations','high citations','very high citations'))
citation_count <- cut(df_citationsreceived$n, breaks = c(0,25,50,75,100,500,max(df_citationsreceived$n)), labels = c('<25 citations', '25-50 citations','50-75 citations','75-100 citations','100-500 citations','more than 500 citations'))
df_citationsreceived$citation_count <- citation_count
ggplot(df_citationsreceived, aes(x = reorder(citation_count,-n, FUN = median),y=n))+geom_boxplot(aes(fill=citation_count))
ggplot(df_citationsreceived%>%dplyr::filter(df_citationsreceived$citation_count!='more than 500 citations'), aes(x = reorder(citation_count,-n, FUN = median),y=n))+geom_boxplot(aes(fill=citation_count))
```
```{r}
df_temp[,'paper_authors_count']<-factor(df_temp[,'paper_authors_count'])
df_temp$num_authors <- forcats::fct_collapse(factor(df_temp$paper_authors_count), 
+                                                 "0"="0", 
+                                                 "1"="1", "2" = "2", "3"="3","4"="4","5"="5",
+                                                 other_level = "5+")
df_temp[,'paper_journal_category']<-factor(df_temp[,'paper_journal_category'])
```
```{r}
df_temp[,'paper_length']<-factor(df_temp[,'paper_length'])
df_temp$num_pages <- forcats::fct_collapse(factor(df_temp$paper_length), 
+                                                 "0"="0", 
+                                                 "1"="1", "2" = "2", "3"="3","4"="4","5"="5",
+                                                 other_level = "5+")
vcd::mosaic(num_pages~num_authors,df_temp,direction=c("v","h"),highlighting_fill = c("blue","lightblue","grey","pink","magenta","purple","yellow"))
```

#References
https://link.springer.com/article/10.1007/s11192-016-1889-2#:~:text=Probably%20some%20factors%20such%20as,and%20discussion%20and%20so%20on.
Category one: “paper related factors”: quality of paper; novelty and interest of subject; characteristics of fields and study topics; methodology; document type; study design; characteristics of results and discussion; use of figures and appendix in papers; characteristics of the titles and abstracts; characteristics of references; length of paper; age of paper; early citation and speed of citation; accessibility and visibility of papers. Category two: “journal related factors”: journal impact factor; language of journal; scope of journal; form of publication. Category three: “author(s) related factors”: number of authors; author’s reputation; author’s academic rank; self-citations; international and national collaboration of authors; authors’ country; gender, age and race of authors; author’s productivity; organizational features; and funding. Probably some factors such as the quality of the paper, journal impact factor, number of authors, visibility and international cooperation are stronger predictors for citations, than authors’ gender, age and race; characteristics of results and discussion and so on.
  
https://www.nature.com/news/papers-with-shorter-titles-get-more-citations-1.18246